\myTitle

### 2023-10-22 body-text-of-paper.tex
Eric edited methods a bit

(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION)

@@

through each non-founder node in a GSP.   Genetic material is segregated through

@@

\item Each genome in a triangle above a founder node is assorted into
two gametes with recombination occurring between each chromosome

@@

a user-specified recombination map.  
\item Each non-founder individual node, $x$, will have exactly two edges coming

@@

\item If a non-founder individual node $x$ has an edge to a sample node,

@@

\item United pairs of gametes (or those remaining after delivery
to a sample node) in each non-founder individual node $x$ are segregated into

@@

genetic material is entirely without replacement so it will not induce any bias in clustering
power assessment; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individual and markers available to

@@

In the gscramble package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses; however, gscramble is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder individual nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other indivduals, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have individual node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of individual nodes and also adjacent to sample-nodes, the necessary conditions 
for a valid GSP are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders|   \\ 
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\ 
\sum_{i=1}^{e_x} G^-_{x,i} + S_x/2 &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfoundse.  \\ 
\end{aligned}
\end{equation}

>>>>

(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION AND MAKE
THE LABELS BIGGER IN THE SQUARES).

@@

through each non-founder node in a GSP\@.   Genetic material is segregated through

@@

\item Each genome in a triangle above a founder node represents a gamete.
These gametes are united within the founder individual and then
following recombination between each chromosome

@@

a user-specified recombination map, 
 one gamete is assorted down each edge from the founder node to its daughter.
\item Each non-founder node, $x$, will have exactly two edges coming

@@

\item If a non-founder node $x$ has an edge to a sample node,

@@

\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into

@@

genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individuals and markers available to

@@

In the gscramble package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses; however, gscramble is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions 
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\ 
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\ 
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\ 
\end{aligned}
\end{equation}
Before simulating from a GSP, gscramble checks to ensure the GSP
satisfies these validity conditions.

\subsection*{Permutation procedures}

gscramble uses a GSP to segregate segments of genetic material from the founders
to the samples.  At the end of segment simulation, the genome of each sample
is recorded as a mosaic of segments that derived from different founders.  In order
to create a data set of markers in each sample, it is necessary only to identify the alleles
that occurred on each segment within the founders and propagate those alleles to the
corresponding segments within the samples.  At the stage of assigning alleles to
founders, gscramble provides a variety of options for to permute alleles amongst
individuals within samples from which the founders are drawn.  By using permutation,
variability in the data is created, but without sampling with replacement, and hence
not inducing RIPI.
